\begin{resumo}[Abstract]
 \begin{otherlanguage*}{english}

The grading of open-ended exam questions is one of the most time-consuming tasks in teaching practice, requiring significant effort and being subject to evaluator subjectivity. This work presents the development and evaluation of an automated essay grading system based on multiple Artificial Intelligence agents, with support for Retrieval-Augmented Generation (RAG) and a divergence-based arbitration mechanism. The system employs two independent grading agents that evaluate each response in parallel, based on a rubric and the course's reference materials, conditionally triggering a third agent (arbiter) when the divergence between scores exceeds a configurable threshold --- a design inspired by Brazil's official ENEM essay grading process. The architecture uses FastAPI for the backend, LangGraph for pipeline orchestration with parallel graders, LangChain for integration with language models (Gemini 2.5 Flash), and ChromaDB for vector storage. The evaluation was conducted through a controlled case study with paired data between experimental conditions, using three open-ended questions on Algorithms and Data Structures and four response quality levels. Results indicate that the system completed the end-to-end flow without failures (100\% structural completeness), exhibited perfect stability at quality extremes and moderate variation for intermediate levels across three repetitions, and demonstrated that RAG positively contributes to grading ($\Delta = +0.11$ overall), with effects concentrated on intermediate and weak responses. The arbitration mechanism was triggered in 17\% to 25\% of cases, stabilizing final scores for ambiguous responses. Score hierarchy reflected quality levels: Excellent (10.00), Intermediate (8.75), Weak (2.12), and Off-topic (0.00). With parallel execution, the system graded 12 responses in 110 seconds (9.2 s/response), representing a reduction of over 97\% compared to manual grading.

\textbf{Keywords:} Automated Grading. Artificial Intelligence. Large Language Models (LLM). Retrieval-Augmented Generation (RAG). Multi-Agent Systems. Educational Assessment.

 \end{otherlanguage*}
\end{resumo}
